Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-09-29 review_responses.tex

\begin{itemize}
  \item \textbf{III. Experimental Setup and Results A-E}: Sections III.A–E now exclusively describe the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests. Each subsection is focused on describing how the experiments are designed, without interpretation of outcomes. For instance, the synthesizer section defines parameter ranges, while the evaluation section specifies how statistical significance is tested. The only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program}, which serves as a transition between results and discussion.

  \item Experimental findings are presented in \textbf{III-F Best Performers for Each Program} and \textbf{III-G Consistency in Rankings}, which together form the \emph{Results} component. Section III-F reports which losses performed best for each synthesizer, while Section III-G quantifies the consistency of rankings across bootstrap resamples. Keeping all numerical findings here avoids overlap with later interpretation.

  \item \textbf{IV. Discussion} then provides the qualitative interpretation, structured into four dedicated subsections:  
    \begin{enumerate}
      \item \textbf{IV-A. Key Findings}: summarizes the central empirical trends, such as the dependency of loss performance on synthesis method.  
      \item \textbf{IV-B. Practical Recommendations}: translates results into concrete recommendations for practitioners and researchers.  
      \item \textbf{IV-C. Illustrating Loss Landscapes}: adds visualizations that explain why certain losses succeed or fail in specific contexts.  
      \item \textbf{IV-D. Applicability to More Complex Synthesizers}: discussion on how observed patterns extend beyond the simplified testbed to richer, real-world synthesizers.  
    \end{enumerate}

  \item To improve clarity, we also relocated explanatory material about experiment design (e.g., parameter settings, optimizer configuration) previously scattered across the results section into Sections III-A–E, ensuring that the results sections remain strictly empirical, while the discussion is reserved for interpretation. 
\end{itemize}

@@

We did not modify the definitions in \textbf{III.A Loss Function Implementation Details}, dedicating this section to only technical explanations.

\subsection*{R1.5 Analysis of Failure or Success.}
\label{R1.5}
\noindent\textbf{Reviewer:} \\
\textbf{Add qualitative or spectral analyses of where specific losses succeed or fail}\\

\noindent\textbf{Response:} \\
This is an interesting suggestion. Our original experiments were designed to emphasize quantitative, statistically robust measurements of loss function performance, which is why we initially avoided focusing on individual examples of sounds, spectrograms, or loss landscapes (in part to reduce the risk of cherry-picking). We do concede that illustrative examples can provide helpful intuition for readers and motivate future research, however, we believe that an analysis of loss landscapes (rather than spectrograms or qualitative tests) is a more useful approach to an explanation for \textit{where} and \textit{why} losses might succeed or fail.

To this end, we have added a new Section \textbf{IV.C Illustrating Loss Landscapes} where qualitative visualizations of loss landscapes  are added. These examples highlight why specific losses may succeed or fail and reinforce the broader importance of loss landscapes in assessing the appropriateness of similarity measures~\cite{vahidi2023mesostructures,turian2020sorry}, supporting our argument that effective navigation of loss landscapes represents an important gap in current knowledge and a possible direction for future work:

@@

\subsection*{R1.6 Applicability to More Complex or Real-world Synthesizers.}
\label{R1.6}
\noindent\textbf{Review:} \\
\noindent \textbf{Expand the discussion to address applicability to more complex or real-world synthesizers}
\\

\noindent\textbf{Response:} \\
We added a new subsection, \textbf{IV.D Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified test bed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings:

\begin{displayquote}
    The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principle with 2 parameters each. While this level of simplicity facilitates controlled comparisons, it does not fully capture the richness of real-world synthesizers, which can have hundreds of parameters with various routes. As a result, the extent of the generalizability of our findings does require further research.

    Nonetheless, we believe that the insights here can extend naturally to more complex domains. Depending on the feature of interest, the observed dependence of loss performance on synthesis method will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.

@@

\noindent\textbf{Response} \\

In \textbf{IV.A Key Findings}, we discuss important findings and provide answers to our research questions. The following section, \textbf{IV.B Practical Recommendations}, distills our findings into actionable heuristics for practitioners.

\begin{displayquote}
    Our findings not only provide answers to the research questions proposed in the introduction, but also serve as guidelines for practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
\begin{itemize}
    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking. However, we encourage confirmation of conclusions with manual listening, particularly if specificity is important. 
    \item For amplitude-modulated synthesis, DTW-based losses such as \DTWEnv{} are recommended over spectrogram differences and JTFS.
    \item  For subtractive/noise-filtered synthesis, spectrogram based losses are most effective. Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
    \item Iterative differentiable optimization via the Faust-to-JAX pipeline is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
    \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}. Thus, our results call into question its effectiveness as a general mesostructure measure, and highlight a need for further research into its utility.
\end{itemize}

\end{displayquote}
\section{Reviewer 2}
\subsection{Original Review}
\noindent
The paper considers the evaluation of different audio similarity metrics for sound-matching applications.  The authors categorize prior methods largely as direct optimization via gradient-free iterative search or open-loop inference via a supervised deep learning model.  Moreover, different approaches which may or may not accommodate out-of-domain sounds.  Accordingly, the authors propose the combined use of differentiable synthesizers and closed-loop iteration as a sound means of achieving synthesizer sound-matching, effectively combining the advantages of otherwise disparate methods (i.e. access to gradients while iteratively improving matching results).  Additionally, within their evaluation of different similarity metrics (parameter losses, spectral distances, etc.), they also introduce a DTW-based similarity metric to partially neutralize the effect of time alignment on similarity scores.  A primary conclusion of the work is that there is no one best sound similarity measure capable of working across all synthesizers.  \\
The subject matter of the work is relevant, the experimental setup is satisfactory, and I do not have reason to doubt the correctness of their analysis.  However, my primary concerns with this work are that:  
\begin{enumerate}
  \item \textbf{(R2.1) It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations that were conducted by the authors.}  
  \item \textbf{(R2.2) The authors claim that differentiable iterative sound-matching is largely unexplored, encouraging development of techniques in this area.}  
    \begin{itemize}
      \item \textbf{Barkan et. al, ``Inversynth II: sound-matching via Self-Supervised Synthesizer-Proxy and Inference-Time Finetuning'' explored precisely this, training an initial ``one-shot'' supervised model, and then performing inference-time optimization, updating model weights to improve the matching against test audio.  Moreover, their use of a synthesizer-proxy negates the explicit need of implementing the synthesizer math differentiably, while still falling under the category of differentiable iterative sound-matching.}  
      \item \textbf{Cherep et. al, ``Creative Text-to-Audio Generation via Synthesizer Programming'' uses a differentiable synthesizer and the CLAP score as a similarity metric for iterative optimization.  While the authors lean into synthesizer sound-matching against a text query, the method would be readily applicable to audio-to-audio sound-matching using the CLAP audio tower alone.}  
    \end{itemize}
  \item \textbf{(R2.3) This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.}
\end{enumerate}

\subsection{R2.1 Clarity of Proposed Method}
\label{R2.1}
\noindent\textbf{Reviewer:}\\
\noindent \textbf{It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations. } 
\\

@@

\end{displayquote}

\medskip
\noindent We restate our methodology in the opening paragraph of the \textbf{III. Experimental Setup} section:

\begin{displayquote}
    At a glance, our methodology is to conduct controlled, low-level differentiable experiments—without proxy networks—by pairing four DSP-based synthesizers with four differentiable loss functions. We optimize parameters directly via gradients in order to isolate loss–synthesizer interactions. This controlled setup systematically evaluates the performance of iterative sound-matching pipelines, where “performance” is defined as the similarity of the synthesized output to a target. Similarity is assessed with P-Loss, MSS, and—most importantly—manual listening tests. This design allows us to directly test our central hypothesis: that the effectiveness of a loss function depends on the synthesis method, and that no universal “best” similarity measure exists.
\end{displayquote}

\subsection{R2.2 Relation to Prior Work}
\label{R2.2}
\noindent\textbf{Reviewer:}\\
\noindent \textbf{Review: The authors claim that differentiable iterative sound-matching is largely unexplored, but prior work (Barkan et al., Cherep et al.) could be considered in this category.}\\

@@

\noindent\textbf{Response:} \\
We thank the reviewer for pointing out sound-matching works that match the ``differentiable and iterative'' description, but were not covered in our literature review. We believe that the term ``under-explored'' is perhaps more appropriate. We've added these works to \textbf{Table 1} which provides a summary of notable previous work, and discussed them in more detail in the final paragraphs of \textbf{II.F Historical Framing of Sound-Matching}: 
\begin{itemize}
  \item \textbf{Barkan et al.:} Their work relies on neural proxies for the synthesizer, with inference-time finetuning of the encoder. This constitutes iterative optimization, but the optimization occurs on an approximate model rather than directly on differentiable DSP functions. By contrast, our approach avoids proxy error and directly analyzes how losses behave relative to known synthesis structures.  
  \item \textbf{Cherep et al.:} Their work uses CLAP embeddings with evolutionary search for text-to-audio tasks. While applicable to audio-to-audio matching, CLAP acts as a black-box representation and does not isolate loss--synth interactions. Our approach is complementary, focusing on controlled experiments with simple differentiable DSP synthesizers to expose the strengths and weaknesses of different similarity measures.  In addition, embedding models such as CLAP use simpler representations at their core---commonly Fourier transformations---and our work explores the low level functions that are often used for the training of embedding models.
\end{itemize}
Looking at these works, we see that our contribution is distinct: we provide the missing \textbf{low-level benchmarking} of loss functions across multiple synthesis methods, which is not addressed by proxy-based or embedding-based approaches.

\subsection*{R2.3 Clear Takeaways}
\noindent\textbf{Reviewer:}\\
\textbf{This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.\\}

\noindent\textbf{Response:} \\
We acknowledge the reviewer’s concern that our main result may appear inconclusive. However, we see this as an important finding in itself: much of the prior literature implicitly over-claims general effectiveness of particular loss functions based on narrow or in-domain settings. Our revisions now explicitly emphasize that the lack of a “silver bullet” for loss functions is a core contribution of this work. By showing that loss effectiveness depends on the synthesis context, the inconclusiveness can be seen as evidence against over-generalization, and an argument for more thorough, context-sensitive benchmarks in measuring sound similarity.\\

@@

As for clear takeaways, we've added two new sections which address the valid issues raised here and by Reviewer 1.

\begin{itemize}
  \item \textbf{IV.B Practical Recommendations} which distills our findings into actionable guidelines for practitioners. DTW_Env is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised). We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.
  \item \textbf{IV.D Applicability to More Complex Synthesizers} that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
\end{itemize}
For a more detailed response, see our response to the Reviewer.1 in Sections~\nameref{R1.6} and \nameref{R1.7}. Our response to \nameref{R1.5} may also be of interest.

@@

\section{Reviewer 4}
Here we've addressed the four bigger concerns first, leaving the minor edits for \nameref{r4.minor}.

\subsection{Original Review}
Well written paper with comprehensive literature review and extensive experiments to bridge existing research gap. Although this paper does not have any ground-breaking new methods that could achieve SOTA performance, the experimental results could still serve as a useful reference for researchers who want to try this field. So I believe it should be accepted with some minor modifications.

(\textbf{R4.1}) Page 3, Left Column, the first equation.
\textbf{I am not sure if `t` is a common symbol to represent `target` in sound marking, but my field of audio/signal processing, `t` usually represents `time`. So it is a bit confusing to me when reading it for the first time.
Would something like $x_\theta$ and $x_t$ be better?} So x always represent the sound. And the subscript indicated if the sound is from the output given the parameter theta, or the target sound. Even better, probably we can use $x_0$ instead of $x_t$ to represent the target sound?

@@

Page 3 Right column section C.
Isn't it better to move the symbol explanation earlier to where the first equation is? I could only under the equations after reading section C.

(\textbf{R4.2}) \textbf{Page 3 Right column at the end
You mentioned "There is no optimal or correct version of an 8-bit snare sound".
This statement is true when you are given only the text description of the sound. However, if you have the target sound $x_0$ (or t in your current draft), the target sound itself is "correct", isn't it?} So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?

@@

Page 6 Right column Section C Point 4.
"out-of-domain sounds is an unexplored area". Unexplored means no one has done it before to me. But in Table 1, you have already cited a few work that tried to do it. So, should it be "underexplored" instead of "unexplored"?

Page 6 Right column Section IV A
You mentioned "four different synthesizers" here, but you never elaborate on which four, not until page 8. Probably you can specify the four synthesizers earlier?

(\textbf{R4.3}) \textbf{Also, in the blinded hearing test, the two listeners are the authors, it seems a bit too biased to me.} Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the authors would yield a more convincing listening results.

@@

Page 7 Left Column 1st Paragraph
You described the Likert scale from 1 to 5 here. But in the next paragraph, you describe the Likert scores again. Not sure if I misunderstood anything, but It looks a bit redundant to me. Would it be possible to combine both to reduce redundancy?

(\textbf{R4.4}) \textbf{Page 7 Right Column section B
"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?}
So the STFT losses doesn't work for the cases where there's time shift?

@@

Page 11, the "weakness" section under Summary
Just a minor grammar question, Is there a missing subject in the first sentence?
While [we] cannot prove ...

(\textbf{R4.5}) O\textbf{ne final question that is just out of curiosity.
In your iterative algorithm, you always run it for 200 steps, right?
In my field of machine learning, if we set 200 steps, the result in the final step might not be the optimal one due to overfitting. So sometimes we obtain the optimal results at around 100 steps, and the model gets worse after this.
Do you have the same thing in your algorithm? Or the final step is always the best step?}

\subsection{\textbf{R4.1} Use of Appropriate Symbols}
\noindent\textbf{Reviewer:}\\
\textbf{The notation \textbf{[for the equation introduced in II.A]} is confusing. In the manuscript, $t$ is used to denote the target sound, but in audio/signal processing $t$ usually represents time.  
A clearer alternative would be to write $x_\theta$ and $x_0$, so that $x$ always denotes a sound waveform, with the subscript indicating whether it is generated from parameters $\theta$ or given as the target.  
This avoids overloading $t$ and improves readability.}
\\

@@

\noindent\textbf{Response:}
This is a helpful suggestion, and we have made the switch from $x$ and $t$ to $x_{\theta}$ and $x_0$ throughout the manuscript.

\subsection{\textbf{R4.2} Optimal Targets}
\noindent\textbf{Reviewer:}\\
\textbf{You mentioned "There is no optimal or correct version of an 8-bit snare sound".
This statement is true when you are given only the text description of the sound. However, if you have the target sound $x_0$  the target sound itself is "correct", isn't it? So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?\\}

@@

\noindent\textbf{Response:}
The original statement referred to how realistically in sound-design, a "correct" solution rarely exists due to our subjective experience of sound. To reduce confusion, the section has been rephrased as follows:
\begin{displayquote}
    Moreover, the ``correct'' solution may not exist in a strict sense: depending on the artistic goal, multiple parameter settings may yield acceptable or even preferable outputs.  In the following sections we will discuss the various components of sound-matching, the subjectivity of ``correct'' solutions, and important issues in the field.
\end{displayquote}

\subsection{R4.3 Hearing Tests}
\label{}
\noindent\textbf{Reviewer:}\\
\textbf{Also, in the blinded hearing test, the two listeners are the authors, it seems a bit too biased to me. Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the author would yield a more convincing listening results.
}
\\

\noindent{\textbf{Response:}}
The listening tests were done without use of any labels or previous exposure to outputs, which we believe minimizes the risk of bias. Unfortunately, adding another listener was not possible given the allotted time for the revisions. If the reviewer feels that this is necessary, it can be done for the next round of reviews.

### 2025-09-29 review_responses.tex

\subsection{\textbf{R4.4} STFT Losses and Time-Shifts}
\noindent\textbf{Reviewer:}\\
\textbf{"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
So the STFT losses doesn't work for the cases where there's time shift?}
\\

@@

\noindent\textbf{Response:}
What we refer to as STFT losses refers to the near ubiquitous approach of applying L1/L2 to STFT spectrograms. This approach does not account for two sounds that are nearly identical but slightly time-shifted~\cite{vahidi2023mesostructures,turian2020sorry}.

\subsection{\textbf{R4.5} Can We Overfit? Is it a Concern?}
\noindent\textbf{Reviewer:}\\
\textbf{One final question that is just out of curiosity.
In your iterative algorithm, you always run it for 200 steps, right?
In my field of machine learning, if we set 200 steps, the result in the final step might not be the optimal one due to overfitting. So sometimes we obtain the optimal results at around 100 steps, and the model gets worse after this.
Do you have the same thing in your algorithm? Or the final step is always the best step?
}
\\

\noindent\textbf{Response:} 
Since we are not training a model that will be applied to other scenarios, overfitting here just means that the parameter updates are stuck around a local or global minimum. This means that \textit{if} the target parameters have been reached, subsequent updates will not diverge from it (or at least stay very close). We can share our visualization of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper. Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates fluctuate slightly and randomly around the target.

@@

\begin{figure*}[h]
    \centering
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/p0_SIMSE_Spec.png}
        \caption{SIMSE\_Spec reaching target and not diverging in \BPNoise}
        \label{fig:p0_simse_iterations}
    \end{minipage}%
    \hfill
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/p2_DTW_Onset.png}
        \caption{DTW\_Env reaching target and not diverging in \AmpMod}
        \label{fig:p2_dtw_iterations}
    \end{minipage}
\end{figure*}

\subsection{R4 Minor Edits}
\label{r4.minor}
\noindent\textbf{Reviewer:}\\
Isn't it better to move the symbol explanation earlier to where the first equation is? I could only under the equations after reading section C.

@@

\noindent\textbf{Response:}
We have moved symbol explanations right after the introduction in 
\textbf{II. A Formalization of Sound-Matching}. 
\\

\noindent\textbf{Reviewer:}\\
"out-of-domain sounds is an unexplored area". Unexplored means no one has done it before to me. But in Table 1, you have already cited a few work that tried to do it. So, should it be "underexplored" instead of "unexplored"?

@@

\noindent\textbf{Response:}
We've changed our phrasing where necessary. See also \nameref{R2.2}.
\\

\noindent\textbf{Reviewer:}\\
You mentioned "four different synthesizers" here, but you never elaborate on which four, not until page 8. Probably you can specify the four synthesizers earlier?

@@

\noindent\textbf{Response:}
We've modified the beginning of Section \textbf{III. Experimental Setup and Results}

\begin{displayquote}
    At a glance, our methodology is to conduct controlled, low-level differentiable experiments without the use of proxy networks. We pair four differentiable losses with four differentiable DSP-based synthesizers: BP-Noise (band-pass noise with LP/HP filters), Add-SineSaw (additive sine and saw oscillators), Noise-AM (noise modulated by an LFO), and SineSaw-AM (sine-modulated saw oscillator). 
\end{displayquote}

\noindent\textbf{Reviewer:}\\
You described the Likert scale from 1 to 5 here. But in the next paragraph, you describe the Likert scores again.

@@

\noindent\textbf{Response:}
This has been fixed in our rewrite of the methodology section (\textbf{III. Experimental Setup and Results}).
\\

\noindent\textbf{Reviewer:}\\
Fix "While [we] cannot prove ..."

### 2025-09-29 main.tex
Updates from Overleaf

\markboth{%
  \parbox{\textwidth}{%
    This work has been submitted to the IEEE for possible publication. \\
    Copyright may be transferred without notice, after which this version may no longer be accessible.%
  }%
}{}

@@

\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}. }

@@

\todo{
Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}. These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values. The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.}

>>>>

% \markboth{%
%   \parbox{\textwidth}{%
%     This work has been submitted to the IEEE for possible publication. \\
%     Copyright may be transferred without notice, after which this version may no longer be accessible.%
%   }%
% }{}

@@

A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.

@@

Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}. These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values. The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.

### 2025-09-29 sections/background_related.tex
Updates from Overleaf

\subsection{Supvervised Versus Direct Optimization}
\label{sec:optimization}

@@

Having looked at the current literature, we identify four major areas of weakness in sound-matching. We target the first two issues in this work, while the latter two issues are left for future work.
\begin{enumerate}
    \item \LossSelect: It remains unclear whether there exists a universally best-performing loss function, 
    or whether performance depends on factors such as the sound domain, synthesizer architecture, 
    and desired output characteristics. Existing studies typically evaluate losses only with simple synthesis 
    setups~\cite{vahidi2023mesostructures}, leaving open questions about their generality. 
    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching. The effects of using different isolated DSP functions in iterative sound-matching and how these interact with loss functions remains largely untested.
    \item \PeriodicLoss: Loss function landscapes are not easy to navigate, as there can be many local minima, or large flat areas~\cite{turian2020sorry,vahidi2023mesostructures}.  In differentiable settings, this can cause gradient descent updates to not reach the global minima. 
    \item \OutDomain: Sound-matching with out-of-domain sounds is a largely unexplored area, yet a necessary one for practical applications for sound designers.
\end{enumerate}

>>>>

\subsection{Supervised Versus Direct Optimization}
\label{sec:optimization}

@@

Having looked at the current literature, we identify four major areas of weakness in sound-matching. We target the first two issues in this work, while the latter two issues are left for future work.
\begin{enumerate}
    \item \LossSelect: It remains unclear whether there exists a universally best-performing loss function, 
    or whether performance depends on factors such as the sound domain, synthesizer architecture, 
    and desired output characteristics. Existing studies typically evaluate losses only with simple synthesis 
    setups~\cite{vahidi2023mesostructures}, leaving open questions about their generality. 
    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching. The effects of using different isolated DSP functions in iterative sound-matching and how these interact with loss functions remains largely untested.
    \item \PeriodicLoss: Loss function landscapes are not easy to navigate, as there can be many local minima, or large flat areas~\cite{turian2020sorry,vahidi2023mesostructures}.  In differentiable settings, this can cause gradient descent updates to not reach the global minima. 
    \item \OutDomain: Sound-matching with out-of-domain sounds is an under-explored area, yet a necessary one for practical applications for sound designers.
\end{enumerate}

### 2025-09-29 sections/introduction.tex
Updates from Overleaf

A digital audio synthesizer is any software used for the creation and manipulation of audio. The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}. A typical synthesizer has a number of parameterizable functions which affect the sound output in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached. The automation of this approach has commonly been referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators ~\cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}. The main requirements of sound-matching are a target sound, a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of the target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}. However, despite the seeming simplicity of the problem and decades of research, sound-matching is not yet a practical solution for sound-designers. Here we are motivated to understand why this is the case, and what future research should prioritize in order to improve the field.

@@

The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}. While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings. Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests? (Q2) Can loss functions based on Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over SOTA loss functions, and under what conditions?
(Q3) Is iterative differentiable optimization a viable strategy for design of sound-matching experiments?

 We adopt a lesser used \textit{iterative} and \textit{differentiable} approach to defining sound-matching experiments. The iterative approach better mimics the manual process of recursive listening and parameter adjustments towards sound design, while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem. We define four differentiable synthesizers (each showcasing a fundamental method of synthesis used in modern synthesizers) and pair them with four different loss functions (two established methods, one utilizing DTW, and one utilizing SIMSE). We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors. We apply statistical ranking applied to the evaluation scores to compare outcomes across synthesizer–loss combinations and measure the similarity of different evaluation measures.

>>>>

A digital audio synthesizer is any software used for the creation and manipulation of audio. The possibilities afforded by these synthesizers are infinite, leading to their widespread adoption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}. A typical synthesizer has a number of parameterizable functions which affect the sound output in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached. The automation of this approach has commonly been referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators ~\cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}. The main requirements of sound-matching are a target sound, a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of the target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}. However, despite the seeming simplicity of the problem and decades of research, sound-matching is not yet a practical solution for sound-designers. Here we are motivated to understand why this is the case, and what future research should prioritize in order to improve the field.

@@

The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}. Testing this hypothesis requires a variety of sound-matching experiments that measure whether a single loss function proves most effective. While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have often been made in limited contexts that may not generalize to other settings. Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests? (Q2) Can loss functions based on Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over SOTA loss functions, and under what conditions?
(Q3) Is iterative differentiable optimization a viable strategy for design of sound-matching experiments?

 We adopt a lesser used \textit{iterative} and \textit{differentiable} approach to defining sound-matching experiments. The iterative approach better mimics the manual process of recursive listening and parameter adjustments towards sound design, while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem. We define four differentiable synthesizers (each showcasing a fundamental method of synthesis used in modern synthesizers) and pair them with four different loss functions (two established methods, one utilizing DTW, and one utilizing SIMSE). We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors. We apply statistical ranking to the evaluation scores to compare outcomes across synthesizer–loss combinations and measure the similarity of different evaluation measures.
